### 2015-02-08 article.Rmd
Added first section, with citations in markdown format.

---
title: "Peer influences in air travel behavior: A social autoregressive analysis"
author: Gregory S. Macfarlane, Thomas A. Wall, Bingling Zhang, and Kari E. Watkins
bibliography: bibliography/bibliography.bib
csl: bibliography/elsevier-harvard.csl 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
---

**Abstract**
Social networks can influence travel in two distinct ways: first, people often make trips to see people they know through work or life (exogenous effects); or they make trips that are similar to those taken by friends or family (endogenous effects). In spite of this, traditional models have not always made use of social network data, because information on social networks is both difficult to collect and challenging to implement in models. In this paper, we appropriate techniques developed for spatial data (specifically, a spatial Durbin count model) to construct a trip generation model for socially linked observations. We train this model on data collected through a web survey that interacts with the Facebook application programming interface (API) to capture respondents’ “friends” lists. We construct a set of “social weights” matrices that captures the social tie-strength between observations. A matrix that weights two observations by the number of their mutual friends provides the maximum model likelihood; this model further suggests that a non-social model may result in erroneous interpretations of the link between demographic characteristics and trip generation.

Introduction
================================================
The goal of a modern activity-based travel model is to accurately represent travel behavior choices on a microscopic level [@Bhat2003]. As researchers have pursued this goal, they have sought to incorporate individual behavioral patterns of various types. For example, researchers have built models that allow different classes of people to place different values on goods [@Walker2002], that capture idiosyncratic preferences for a particular lifestyle [@Cao2007a], or that allow multiple decisions to be made either sequentially or simultaneously [@Bhat2006; @Pinjari2011].

Another dimension of individual behavior that is receiving increasing attention from transportation researchers is the role that peer or social networks play in travel choices. These influences may come in several ways:

  - *Direct*: each individual chooses activities or destinations where their social connections are, or where they will be.
  - *Indirect*: an individual’s social connections influence their choice of destinations or activities.
  - *Endogenous*: individuals construct their social networks with others of shared interest or experience.

The distinction between these elements can be subtle. Imagine two individualswho are socially connected to each other. If these individuals meet for dinner, this is a direct influence, because each person’s choice of activity and destination is dependent on the other person being there. On the other hand if one recommends a restaurant to the other, this is an indirect influence because the restaurant entered one’s choice set through the other’s experience and not his own. Finally, if the two know each other through another shared destination or interest (e.g. workplace, neighborhood, church), the restaurant in question may have entered their choice set by being near one of these destinations.

These elements are important to capture in activity-based models for several reasons. First, they are a natural element of every day behavior. But more importantly, an activity model that does not consider them might produce spurious or unreliable predictions, because standard econometric assumptions that permit consistency and validity may not apply. Despite this importance, models that incorporate social influences are rare for two reasons:  because social network data has been difficult to collect, and because there has not been a convenient econometric framework to apply.

However, both of these situations are changing. Usage of online social networking sites and other social media continues to grow, and contains social connection and preference data for  millions of individuals. On the econometric front, the spatial econometrics literature in the last twenty-plus years [@Anselin2003; @Anselin2011] enables the view of spatial interaction as a special case of social interaction, or as a proxy that may be more easily observed in data.

In this paper, we develop and apply a technique that uses spatial data to analyze airline travel (trip frequency)  of individuals in a common social network. To do this, we construct a series of social weights matrices representing potential relationship strength between members of this network, and incorporate these matrices in a social Durbin model capturing direct, indirect, and endogenous effects. 

The paper is organized as follows: Section 2 presents the literature establishing spatial/social econometrics as well as previous studies establishing the effect of social networks on travel behavior. Section 3 presents our model and the training data that we collected from a sample of undergraduate students (specifically those in the fraternity and sorority, or "Greek," system), with results from the model presented in Section 4. Section 5 concludes the paper with implications for future practice and research.

### 2015-02-18 article.Rmd
Limited line length in paragraphs.

@@
 **Abstract**
-Social networks can influence travel in two distinct ways: first, people often make trips to see people they know through work or life (exogenous effects); or they make trips that are similar to those taken by friends or family (endogenous effects).
-In spite of this, traditional models have not always made use of social network data, because information on social networks is both difficult to collect and challenging to implement in models.
-In this paper, we appropriate techniques developed for spatial data (specifically, a spatial Durbin count model) to construct a trip generation model for socially linked observations.
-We train this model on data collected through a web survey that interacts with the Facebook application programming interface (API) to capture respondents’ “friends” lists.
-We construct a set of “social weights” matrices that captures the social tie-strength between observations.
-A matrix that weights two observations by the number of their mutual friends provides the maximum model likelihood; this model further suggests that a non-social model may result in erroneous interpretations of the link between demographic characteristics and trip generation.
+Social networks can influence travel in two distinct ways: first, people often
+make trips to see people they know through work or life (exogenous effects); or
+they make trips that are similar to those taken by friends or family (endogenous
+effects). In spite of this, traditional models have not always made use of
+social network data, because information on social networks is both difficult to
+collect and challenging to implement in models. In this paper, we appropriate
+techniques developed for spatial data (specifically, a spatial Durbin count
+model) to construct a trip generation model for socially linked observations. We
+train this model on data collected through a web survey that interacts with the
+Facebook application programming interface (API) to capture respondents’
+“friends” lists. We construct a set of “social weights” matrices that captures
+the social tie-strength between observations. A matrix that weights two
+observations by the number of their mutual friends provides the maximum model
+likelihood; this model further suggests that a non-social model may result in
+erroneous interpretations of the link between demographic characteristics and
+trip generation.
 
@@
 ================================================
-The goal of a modern activity-based travel model is to accurately represent travel behavior choices on a microscopic level [@Bhat2003].
-As researchers have pursued this goal, they have sought to incorporate individual behavioral patterns of various types.
-For example, researchers have built models that allow different classes of people to place different values on goods [@Walker2002], that capture idiosyncratic preferences for a particular lifestyle [@Cao2007a], or that allow multiple decisions to be made either sequentially or simultaneously [@Bhat2006; @Pinjari2011].
+The goal of a modern activity-based travel model is to accurately represent
+travel behavior choices on a microscopic level [@Bhat2003]. As researchers
+have pursued this goal, they have sought to incorporate individual behavioral
+patterns of various types. For example, researchers have built models that allow
+different classes of people to place different values on goods [@Walker2002],
+that capture idiosyncratic preferences for a particular lifestyle [@Cao2007a],
+or that allow multiple decisions to be made either sequentially or
+simultaneously [@Bhat2006; @Pinjari2011].
 
-Another dimension of individual behavior that is receiving increasing attention from transportation researchers is the role that peer or social networks play in travel choices.
-These influences may come in several ways:
+Another dimension of individual behavior that is receiving increasing attention
+from transportation researchers is the role that peer or social networks play in
+travel choices. These influences may come in several ways:
 
-  - *Direct*: each individual chooses activities or destinations where their social connections are, or where they will be.
-  - *Indirect*: an individual’s social connections influence their choice of destinations or activities.
-  - *Endogenous*: individuals construct their social networks with others of shared interest or experience.
+	- *Direct*: each individual chooses activities or destinations where their
+	  social connections are, or where they will be.
+	- *Indirect*: an individual’s social connections influence their choice of
+	  destinations or activities.
+	- *Endogenous*: individuals construct their social networks with others of
+	  shared interest or experience.
 
-The distinction between these elements can be subtle.
-Imagine two individualswho are socially connected to each other.
-If these individuals meet for dinner, this is a direct influence, because each person’s choice of activity and destination is dependent on the other person being there.
-On the other hand if one recommends a restaurant to the other, this is an indirect influence because the restaurant entered one’s choice set through the other’s experience and not his own.
-Finally, if the two know each other through another shared destination or interest (e.g.
-workplace, neighborhood, church), the restaurant in question may have entered their choice set by being near one of these destinations.
+The distinction between these elements can be subtle. Imagine two individuals who
+are socially connected to each other. If these individuals meet for dinner, this
+is a direct influence, because each person's choice of activity and destination
+is dependent on the other person being there. On the other hand if one
+recommends a restaurant to the other, this is an indirect influence because the
+restaurant entered one's choice set through the other's experience and not his
+own. Finally, if the two know each other through another shared destination or
+interest (e.g. workplace, neighborhood, church), the restaurant in question may
+have entered their choice set by being near one of these destinations.
 
-These elements are important to capture in activity-based models for several reasons.
-First, they are a natural element of every day behavior.
-But more importantly, an activity model that does not consider them might produce spurious or unreliable predictions, because standard econometric assumptions that permit consistency and validity may not apply.
-Despite this importance, models that incorporate social influences are rare for two reasons:  because social network data has been difficult to collect, and because there has not been a convenient econometric framework to apply.
+These elements are important to capture in activity-based models for several
+reasons. First, they are a natural element of every day behavior. But more
+importantly, an activity model that does not consider them might produce
+spurious or unreliable predictions, because standard econometric assumptions
+that permit consistency and validity may not apply. Despite this importance,
+models that incorporate social influences are rare for two reasons:  because
+social network data has been difficult to collect, and because there has not
+been a convenient econometric framework to apply.
 
-However, both of these situations are changing.
-Usage of online social networking sites and other social media continues to grow, and contains social connection and preference data for  millions of individuals.
-On the econometric front, the spatial econometrics literature in the last twenty-plus years [@Anselin2003; @Anselin2011] enables the view of spatial interaction as a special case of social interaction, or as a proxy that may be more easily observed in data.
+However, both of these situations are changing. Usage of online social
+networking sites and other social media continues to grow, and contains social
+connection and preference data for  millions of individuals. On the econometric
+front, the spatial econometrics literature in the last twenty-plus years
+[@Anselin2003; @Anselin2011] enables the view of spatial interaction as
+a special case of social interaction, or as a proxy that may be more easily
+observed in data.
 
-In this paper, we develop and apply a technique that uses spatial data to analyze airline travel (trip frequency)  of individuals in a common social network.
-To do this, we construct a series of social weights matrices representing potential relationship strength between members of this network, and incorporate these matrices in a social Durbin model capturing direct, indirect, and endogenous effects.
-
+In this paper, we develop and apply a technique that uses spatial data to
+analyze airline travel (trip frequency)  of individuals in a common social
+network. To do this, we construct a series of social weights matrices
+representing potential relationship strength between members of this network,
+and incorporate these matrices in a social Durbin model capturing direct,
+indirect, and endogenous effects.
 
-The paper is organized as follows: Section 2 presents the literature establishing spatial/social econometrics as well as previous studies establishing the effect of social networks on travel behavior.
-Section 3 presents our model and the training data that we collected from a sample of undergraduate students (specifically those in the fraternity and sorority, or "Greek," system), with results from the model presented in Section 4.
-Section 5 concludes the paper with implications for future practice and research.
+The paper is organized as follows: Section 2 presents the literature
+establishing spatial/social econometrics as well as previous studies
+establishing the effect of social networks on travel behavior. Section 3
+presents our model and the training data that we collected from a sample of
+undergraduate students (specifically those in the fraternity and sorority, or
+"Greek," system), with results from the model presented in Section 4. Section 5
+concludes the paper with implications for future practice and research.
+
+Social Networks and Social Econometrics
+=======================================================
+
+Social Networks, Online Observation, and Facebook
+-------------------------------------------------------
+The concept of a social network provides a construct in which the structural and
+functional dimensions of individuals’ social relationships (O'Reilly 1988)
+can be systematically examined for social influence, or other characteristics of
+interest.  Social network structure, which refers to the individuals and
+linkages comprising one's social network (Due, Holstein et al. 1999) can be
+analyzed either through egocentric network analysis (focused on a single
+individual and his relationships with others) or whole network analysis (focused
+comprehensively on the relationships among all members in a network).  Social
+network function clarifies the nature of, and interpersonal interaction among,
+the linkages within the social network structure; for example, "social support,
+relational strain, or social anchorage," among others. (Due, Holstein et al.
+1999) In the context of this study, one important dimension of social network
+function could be the level of influence associated with an individual
+connection between two individuals, or among many connections within a group of
+individuals.
+
+Data collection, however, complicates this type of study.   The predominant
+method employed by previous studies (for example, (27–29)) to construct
+models of individuals’ social networks, called “name generators,” uses sets of
+interview questions designed to solicit information from an individual regarding
+whom he considers as part of his social network (structure) and the role or
+importance of that individual (function).  The name generator approach can
+become problematic, however, when respondents have difficulty in recalling
+information regarding their social networks (15), thus providing information
+to the researchers that is unknowingly incomplete.  In addition, the
+name-generator approach can be  time consuming, with data collection interviews
+frequently taking several hours for a single interview (30) to construct a
+model of one individual's egocentric social network.
+
+Online social networking websites, such as Facebook and Twitter, provide an
+alternative source for social network data, the collection of which allows for
+the delineation of naturally developing social networks with unambiguous
+boundaries (thus overcoming respondent recollection issues) (14, 15).
+Additionally, as such data is stored in a centralized computer database and in a
+format that is consistent across all users, social network data can be collected
+in seconds and with greater structural consistency.  Thus, data collection is
+far less time consuming and data quality and consistency is potentially improved
+as compared to the name generator approach (14, 15).
+
+In this study, online social network data collection is focused exclusively on
+users of the social networking site, Facebook. With 864 million daily active
+users, and 1.12 billion monthly active users globally as of September 2014,
+Facebook is among the world's largest social networking sites (Facebook
+2015) (by comparison, Twitter has 284 million monthly active users across the
+globe as of October 2014). (Twitter 2015).  Facebook enables users to create
+online profiles (e.g., general demographic, educational, special interest, and
+geographical information), which is shared with other users.  As users connect
+with other individuals, becoming 'friends,' an online model of social
+relationships and connections is created. (Wall, Macfarlane et al. 2014). A
+study by the Pew Research Center found that the primary motivation for the
+majority of adult  users of social networking sites in the United States is
+"staying in touch with family members," and "staying in touch with current
+friends and reconnecting with old friends." (Smith 2011) It therefore may be
+reasonable to "suspect that an individual’s online social network structure
+represents, to some extent, an online model of that individual’s real-life
+social network structure." (Wall, Macfarlane et al. 2014)
+
+There are also several characteristics of the Facebook dataset, as well as the
+types of data contained in that dataset, that are particularly relevant to the
+consideration of socio-spatial influences. Lewis et al. (2008) discusses
+five defining features of the Facebook social network model:
+
+	1. Naturally Occurring:  Facebook data is naturally occurring when downloaded
+     directly from Facebook (i.e. not subject to a respondent recollection ability) 
+	2. Whole Networks: Facebook.com contains complete network data of a
+     sociocentric nature, as opposed to an egocentric nature, which reflects the
+     interconnectedness of network members
+	3. Longitudinal Data: Facebook data may reflect how relationships grow and
+     change over time; 
+	4. Cultural Insight: Facebook users can indicate tastes and values, which may
+     be reflected in the data 
+	5. Relationship Strength: Facebook can provide data on relationship strength.
+
+Several of these items, as well as the types of data contained in Facebook
+datasets are particularly relevant to the consideration of socio-spatial
+influences.
+
+The whole network aspect of Facebook data enables an analysis of interconnected
+or shared relationships among a population.  Further, relationship strength
+provides some insight into the function of relationships, as may be related to
+the level of indirect influence across a relationship.  Lastly, and very
+importantly, Facebook maintains locational information that can easily be
+translated to geospatial
+
+Spatial Econometrics
+---------------------------------------
+Spatial econometrics is built on the understanding that observations near to
+each other spatially are likely to exhibit correlation and interdependence, as
+they share common unobserved variables and influence each other. A class of
+models developed to solve these problems was pioneered by Anselin (1988);
+LeSage & Pace (2009) provide a full and modern treatment.
+
+For a linear regression model , a necessary assumption is that the observations
+are independent, which is unlikely in spatial data. Tobler’s law implies that if
+two observations are “close” to each other, then they are likely to share common
+characteristics (Tobler 1970). Further, observations close to each other may
+actually exert an influence on each other in the way air pollution lowers the
+price of nearby properties (Kim & Goldsmith 2008). If correlation and
+interdependence are not controlled in an econometric model, then the model may
+return biased parameters and/or invalid hypothesis tests.
+
+This correlation and interdependence can be accommodated with a spatial Durbin
+model (SDM),
+
+where  is a spatial weights matrix in which each element  maps the spatial
+relationship between two observations  and ; if  and  are “neighbors,” then .
+The parameter  measures the dependence between  and , and the lagged independent
+variable coefficients in  contain the impact that attributes of  have on .
+
+Such spatial models have been applied to numerous different contexts, including
+transportation modeling and forecasting. In most if not all of these cases,
+however, spatial data is used as a proxy for an inferred social relationship.
+For example, Adjemian et al. (2010) observed that vehicle body types were
+spatially autocorrelated, with the authors inferring that neighbors may
+influence each other in choosing vehicle types. Similarly, Bhat et al.
+(2010) used spatial proximity as a means to study the social destinations of
+teenagers. In these and other cases, it may be more appropriate to use the
+social network directly, if the social network can be translated to a  matrix.
+This is effectively the approach of Bramoullé et al. (2009) and
+Calvó-Armengol et al, (2009), who showed that a student's school performance
+and physical activity level (respectively) are socially interdependent: students
+with studious or athletic friends tend to be so as well.
+
+There are two primary difficulties in creating , one econometric and one
+practical. From an econometric standpoint, the researcher typically must specify
+exogenously. Though some common practices have emerged in the spatial literature
+(Dubin 1998), not all of these can be readily adopted to social networks.
+This gap in the literature is one that we seek to fill with this study, by
+examining several possible  matrices that we construct from social network data.
+The practical difficulty is actually collecting that data, though this is
+becoming less difficult as interaction within social networks happens through
+online sites such as Facebook. We discuss this practical evolution in Section
+2.2.
+
+Methodology
+======================================
+We introduce social dependence into a Poisson count model that predicts the
+number of airline trips an individual makes as a function of his or her
+socioeconomic variables. The Poisson count model estimates the effect of
+dependent variables  on the conditional mean number of trips, . We consider
+social interaction in this model by suggesting that  (where  is the row of  that
+contains the weights for the neighbors of ). Our goal in this analysis is to
+recover a weighting scheme  that maximizes the model likelihood, and use this
+model to examine the inferential error that could result from ignoring social
+interdependence and correlation in such model.
+
+Data Collection
+----------------------------------------
+To examine the potential for influences to travel behavior between individuals’
+within an interconnected social network, a web-based survey was developed, and
+respondents were recruited through the Georgia Tech fraternity and sorority
+system (a social network that may presumably have a proportionally higher level
+of interconnectedness), in partnership with the Georgia Tech Dean of Students,
+Office of Greek Affairs.  The web-based survey instrument was distributed via
+email to the primary points of contact for each fraternity and sorority on file
+at the Office of Greek Affairs.  Fraternities and sororities were asked to also
+encouraged to solicit responses from their alumni, thus, potentially increasing
+the sample population of approximatley 3,300 students in Greek housing.
+Fraternities and sororities were incentivized with the promise of a $500
+charitable contribution made on behalf of the fraternity and the sorority each
+with the highest participation.
+
+The online survey instrument was adapted from Wall et al (2014), and
+contains three principal components.  For a detailed description of the
+variables collected, refer to Wall et al. (2014).  The first component
+collected respondents’ air travel information from the past year – this includes
+a list of trips taken with origin and destination airports for round-trip,
+multi-destination, and one-way travel.  The second component collected
+socio-demographic information, including age, gender, income, education level,
+citizenship, occupation, and race.  The third component asked respondents to log
+in to their Facebook accounts, and the survey automatically collected data
+contained in the respondent's profile and friends list via the Facebook
+application programming interface (API).  This data included the hometown city
+and current city locations of the respondents, and also that of the respondents’
+friends.  The program also collected an encrypted (one way hash) version of the
+Facebook ID for all respondents and their friends.  This ensured subject
+anonymity while simultaneously allowing observations regarding 1) which
+respondents were friends with each other, and 2) how many friends respondents
+have in common, 3) identifying and limiting the potential for duplicate
+respondents.
+
+Because the survey program collects Facebook profile and friends list data from
+the perspective of the respondent's individual profile, it is limited in its
+data collection to "visible" user profile and friend data. That is, individuals'
+information that was restricted as a result of more stringent individual privacy
+settings would not be visible to the survey program. A consequence of this is
+potential for imbalance in privacy settings (and thus, data visibility) between
+two connections, is that friend connections may be  observed as one-way in the
+data even though one-way friend connections are not possible on Facebook.
+
+Facebook locational data (i.e., hometown and current city location) is not
+geocoded, and therefore required some post-processing prior to any analyses.
+Facebook location data were post-processed using R statistical programming
+language and ArcGIS 10.  The study analyzed respondents’ friends who do not
+consider Atlanta, Georgia as their hometown.  Friends’ hometown cities were
+geocoded as point locations. To associate cities as within (or outside of) the
+Atlanta metropolitan region, circular buffers were created (using
+Hartsfield-Jackson Atlanta International airport as the centroid) at 25 mile
+radius increments from 25 miles to 150 miles. The 6 resulting hometown buffer
+regions were then associated with the geocoded city locations, and thus the
+number of friends with hometown city locations within each buffer region were
+determined. Only data from respondents that completed the entire survey (i.e.
+the air travel diary, socio-demographic section, and the Facebook data
+collection) were considered for the analysis.
+
+Candidate Weighting Schema
+-----------------------------------------
+The social network data we retrieved from Facebook allow us to construct four
+social weights matrices, which we describe in this section. To aid this
+presentation, we must first establish some basic notation following elementary
+set theory. The set of friends of an individual  is defined as . The cardinality
+of this set is the number of friends, so therefore the quantity  represents the
+number of friends that individuals  and  have in common.
+
+**Direct Friendship**
+This schema defines a simple binary matrix where a cell is nonzero if  and  are
+friends. Specifically,
+
+**Mutual Friendship**
+This schema defines a matrix where the cell value is the number of friends  and
+have in common (as in Gilbert & Karahalios 2009; He et al. 2012),
+
+**Network Overlap**
+This schema defines a matrix where the cell value is the relative overlap of
+each  pair’s social network (as in Adamic & Adar 2003),
+
+**Direct Network Overlap**
+This schema is similar to the Network Overlap, except that we give a bonus to
+pairs that are themselves friends,
+
+By Facebook’s internal logic, each of these matrices must be symmetric: it is
+not possible on Facebook to have non-reciprocal friends. However, the individual
+user privacy settings make it possible that we would observe apparent one-way
+relationships through the API; indeed, this affected approximately 10% of cases.
+We therefore coerce this and other schema to be symmetric as appropriate.
+
+As a final note on the schema development, behavioral inference on and
+estimation of the models is simplified by using a row-standardized matrix, where
+the rows sum to unity(LeSage & Pace 2009, p.85). We follow this convention
+in all cases.
+
+Results and Discussion
+================================
+
+Conclusions
+================================
+
+# References {-}

### 2015-02-18 article.Rmd
Fixed formatting of lists and added real citations to econometrics section.

@@
 they share common unobserved variables and influence each other. A class of
-models developed to solve these problems was pioneered by Anselin (1988);
-LeSage & Pace (2009) provide a full and modern treatment.
+models developed to solve these problems was pioneered by @Anselin1988;
+@LeSage2009 provide a full and modern treatment.
 
@@
 two observations are “close” to each other, then they are likely to share common
-characteristics (Tobler 1970). Further, observations close to each other may
+characteristics [@Tobler1970]. Further, observations close to each other may
 actually exert an influence on each other in the way air pollution lowers the
-price of nearby properties (Kim & Goldsmith 2008). If correlation and
+price of nearby properties [@Kim2008]. If correlation and
 interdependence are not controlled in an econometric model, then the model may
@@
 model (SDM),
-
-where  is a spatial weights matrix in which each element  maps the spatial
-relationship between two observations  and ; if  and  are “neighbors,” then .
-The parameter  measures the dependence between  and , and the lagged independent
-variable coefficients in  contain the impact that attributes of  have on .
+$$y = \rho W y + X\beta + WX\gamma + \epsilon$$
+where $W$ is a spatial weights matrix in which each element  maps the spatial
+relationship between two observations $i$ and $j$; if $i$ and $j$ are
+"neighbors," then $w_{ij} > 0$. The parameter $\rho$ measures the dependence
+between $y_i$ and $y_j$, and the lagged independent variable coefficients in
+$\gamma$ contain the impact that attributes of $j$ have on $y_i$.
 
@@
 however, spatial data is used as a proxy for an inferred social relationship.
-For example, Adjemian et al. (2010) observed that vehicle body types were
+For example, @Adjemian2010 observed that vehicle body types were
 spatially autocorrelated, with the authors inferring that neighbors may
-influence each other in choosing vehicle types. Similarly, Bhat et al.
-(2010) used spatial proximity as a means to study the social destinations of
-teenagers. In these and other cases, it may be more appropriate to use the
-social network directly, if the social network can be translated to a  matrix.
-This is effectively the approach of Bramoullé et al. (2009) and
-Calvó-Armengol et al, (2009), who showed that a student's school performance
-and physical activity level (respectively) are socially interdependent: students
-with studious or athletic friends tend to be so as well.
+influence each other in choosing vehicle types. Similarly, @Bhat2010b used
+spatial proximity as a means to study the social destinations of
+teenagers, assuming that teenagers who live more closely to each other are more
+likely to be in each other's social networks. In these and other cases, it may
+be more appropriate to use the social network directly, if the social network
+can be translated to a $W$ matrix. This is effectively the approach of
+@Bramoulle2009 and @Calvo-Armengol2009, who showed that a student's school
+performance and physical activity level (respectively) are socially
+interdependent: students with studious or athletic friends tend to be so as
+well.
 
@@
 exogenously. Though some common practices have emerged in the spatial literature
-(Dubin 1998), not all of these can be readily adopted to social networks.
+[@Dubin1998], not all of these can be readily adopted to social networks.
 This gap in the literature is one that we seek to fill with this study, by

### 2015-02-19 rscripts/datacleaner.R
Started work on data cleaning script

@@ -1,0 +1,32 @@
+# Load and do basic data munging on fraternity surveys.
+library(dplyr)
+library(data.table)
+library(igraph)
+
+# Load user data and trim to useable fields
+user <- read.csv("../data/PRIVATEDATA/user.csv", stringsAsFactors = FALSE) %>%
+  select(
+    id, fb_id, zip, location_from, lives_in, age, gender, race,
+    friend_list_size, collected_friends_size
+    ) %>%
+  setnames(
+    c("id", "fbid", "zip", "from", "at", "age", "sex", "race", "friends", 
+      "collected")
+    ) 
+
+# useable records must have friends collected
+user <- user %>%
+  mutate_each(funs(ifelse( . == "NULL", NA, .))) %>%
+  filter(collected > 0) 
+
+# make sure numeric cols are num, character are chr
+user <- user %>%
+  mutate_each(funs(as.numeric(.)), age, friends, collected) %>%
+  mutate(zip = sprintf("%05d", zip))
+
+
+# calculate trips for each user
+
+
+
+

### 2015-03-20 article.Rmd
Change from Rda to rds saving format.

@@
 Facebook ID for all respondents and their friends.  This ensured subject
-anonymity while simultaneously allowing observations regarding 1) which
-respondents were friends with each other, and 2) how many friends respondents
-have in common, 3) identifying and limiting the potential for duplicate
-respondents.
+anonymity while simultaneously allowing us to observe which
+respondents were friends with each other, as well as how many friends respondents
+have in common even if those common friends did not participate in the survey.
+
+```{r load_user}
+user <- readRDS("data/clean/user.rds")
+trips <- readRDS("data/clean/trips.rds")
+friends <- readRDS("data/clean/friends.rds")
+```
 

### 2015-03-20 rscripts/datacleaner.R

@@ -24,9 +24,11 @@ user <- user %>%
   mutate_each(funs(as.numeric(.)), age, friends, collected) %>%
   mutate(zip = sprintf("%05d", zip))
 
+saveRDS(user, file = "../data/clean/user.rds")
 
 # calculate trips for each user
 
 
+saveRDS(friends, file = "../data/clean/friends.rds")
 
 

### 2015-03-20 article.Rmd
Improved text in econometrics section.

@@
 
-For a linear regression model , a necessary assumption is that the observations
-are independent, which is unlikely in spatial data. Tobler’s law implies that if
+For a linear regression model $y = X\beta + epsilon$, a necessary assumption is 
+that the observations are independent, which is unlikely in spatial data. 
+Tobler’s law suggests that if
 two observations are “close” to each other, then they are likely to share common
@@
 
-There are two primary difficulties in creating , one econometric and one
-practical. From an econometric standpoint, the researcher typically must specify
+There are two primary difficulties in creating $W$, one econometric and one
+practical. From an econometric standpoint, the researcher typically must specify $W$
 exogenously. Though some common practices have emerged in the spatial literature
@@
 examining several possible matrices that we construct from social network data.
-The practical difficulty is actually collecting that data, though this is
+The practical difficulty is actually collecting such data, though this is
 becoming less difficult as interaction within social networks happens through
-online sites such as Facebook. We discuss this practical evolution in Section
-2.2.
+online sites such as Facebook. We discussed this practical evolution in Section
+2.1.
 
@@
 
-The online survey instrument was adapted from Wall et al (2014), and
-contains three principal components.  For a detailed description of the
-variables collected, refer to Wall et al. (2014).  The first component
+The online survey instrument was adapted from @Wall2014; a detailed description 
+of the instrument can be found there. The survey consists of three principal 
+components. The first component
 collected respondents’ air travel information from the past year – this includes

### 2015-03-20 article.Rmd
Activated the voice in the survey discussion.

@@
 To examine the potential for influences to travel behavior between individuals’
-within an interconnected social network, a web-based survey was developed, and
-respondents were recruited through the Georgia Tech fraternity and sorority
+within an interconnected social network, we developed a web-based survey 
+and recruited respondents through the Georgia Tech fraternity and sorority
 system (a social network that may presumably have a proportionally higher level
-of interconnectedness), in partnership with the Georgia Tech Dean of Students,
-Office of Greek Affairs.  The web-based survey instrument was distributed via
-email to the primary points of contact for each fraternity and sorority on file
-at the Office of Greek Affairs.  Fraternities and sororities were asked to also
-encouraged to solicit responses from their alumni, thus, potentially increasing
-the sample population of approximatley 3,300 students in Greek housing.
-Fraternities and sororities were incentivized with the promise of a $500
-charitable contribution made on behalf of the fraternity and the sorority each
-with the highest participation.
+of interconnectedness than the general student population), in partnership with 
+the Georgia Tech Dean of Students, Office of Greek Affairs. We distributed the
+web-based survey instrument via email to the primary points of contact for each
+fraternity and sorority on file at the Office of Greek Affairs. To increase the 
+sampling frame potentially beyond the 3,300 currrent fraternity and sorority 
+members, we encouraged the organizations to solicit responses from their alumni
+groups. The fraternity and sorority with the highest response volumes each
+had a $500 donation made in their name to a national charitable organization.
 
@@
 ```{r load_user}
+trips <- readRDS("data/clean/trips.rds")
 user <- readRDS("data/clean/user.rds")
-trips <- readRDS("data/clean/trips.rds")
 friends <- readRDS("data/clean/friends.rds")
@@
 settings would not be visible to the survey program. A consequence of this is
-potential for imbalance in privacy settings (and thus, data visibility) between
-two connections, is that friend connections may be observed as one-way in the
-data even though one-way friend connections are not possible on Facebook.
+that one-way friendships, may be observed in the data even though one-way 
+friend connections are not possible on Facebook.
 

### 2015-03-27 article.Rmd
Added code to calculate matrix weights.

@@
 ---
+```{r setup}
+library(knitr) # nice output tables
+
+# knit options
+opts_chunk$set(
+  cache = TRUE, cache.path = "./.cache/",
+  echo = TRUE, 
+  fig.keep = TRUE, fig.path = "./.figs/"
+)
+
+library(dplyr, warn.conflicts = FALSE) # super data wrangling
+library(igraph) # graph analysis
+library(networkD3) # interactive graph visualizer
+```
 
@@
 number of friends that individuals  and  have in common.
+```{r connections_frame}
+friend_connections <- tbl_df(expand.grid(id_i = user$id, id_j = user$id)) %>%
+  group_by(id_i, id_j)
+```
 
@@
 
+$$ w_{ij} =
+  \begin{cases}
+   1 & \text{if } i \in \phi_j \\
+   0 & \text{if } i \notin \phi_j
+  \end{cases} $$
+
+```{r direct_friends_graph}
+check_friends <- function(id_i, id_j, user, friends){
+  sum(user$fbid[user$id == id_i] %in% friends$fbid[friends$id == id_j])
+}
+
+friend_connections <- friend_connections %>%
+  mutate(direct = check_friends(id_i, id_j, user, friends))
+```
 **Mutual Friendship**
@@
 have in common (as in Gilbert & Karahalios 2009; He et al. 2012),
+```{r mutual_friends}
+count_mutual_friends <- function(id_i, id_j, friends){
+  length(intersect(
+  # sets of friend ids
+    friends$fbid[friends$id == id_i],
+    friends$fbid[friends$id == id_j]
+  ))
+}
+
+friend_connections <- friend_connections %>%
+  mutate(mutual = count_mutual_friends(id_i, id_j, friends))
+```
 
@@
 each  pair’s social network (as in Adamic & Adar 2003),
+```{r network_overlap}
+calculate_newtork_overlap <- function(id_i, id_j, friends){
+  mutual <- length(intersect( 
+    # sets of friend ids
+    friends$fbid[friends$id == id_i],
+    friends$fbid[friends$id == id_j]
+  ))
+
+  total <- length(union( 
+    friends$fbid[friends$id == id_i],
+    friends$fbid[friends$id == id_j]
+  ))
+
+  mutual/total
+}
+
+friend_connections <- friend_connections %>%
+  mutate(overlap = calculate_newtork_overlap(id_i, id_j, friends))
+```
 
@@
 pairs that are themselves friends,
+```{r direct_network_overlap}
+friend_connections <- friend_connections %>%
+  mutate( direct_overlap = overlap + direct )
+```
 

### 2015-03-27 article.Rmd
Added mathematical descriptions of matrices.

@@
 -----------------------------------------
-The social network data we retrieved from Facebook allow us to construct four
-social weights matrices, which we describe in this section. To aid this
-presentation, we must first establish some basic notation following elementary
-set theory. The set of friends of an individual  is defined as . The cardinality
-of this set is the number of friends, so therefore the quantity  represents the
-number of friends that individuals  and  have in common.
+The social network data we retrieved from Facebook allow us to construct four social weights matrices, which we describe in this section.
+To aid this presentation, we must first establish some basic notation following elementary set theory.
+The set of friends of an individual $i$ is defined as $\phi_i$.
+The cardinality of this set is the number of friends, so therefore the quantity $|\phi_i \cup \phi_j|$ represents the number of friends that individuals $i$ and $j$ have in common.
+
 ```{r connections_frame}
@@
 ```
+
 **Mutual Friendship**
 This schema defines a matrix where the cell value is the number of friends  and
-have in common (as in Gilbert & Karahalios 2009; He et al. 2012),
+have in common (as in Gilbert & Karahalios 2009; He et al. 2012),
+
+$$ w_{ij} = |\phi_i \cup \phi_j| $$
+
 ```{r mutual_friends}
@@
 This schema defines a matrix where the cell value is the relative overlap of
-each  pair’s social network (as in Adamic & Adar 2003),
+each pair’s social network (as in Adamic & Adar 2003),
+
+$$ w_{ij} = \frac{|\phi_i \cup \phi_j|} {|\phi_i \cap \phi_j|} $$
+
 ```{r network_overlap}
@@
 **Direct Network Overlap**
-This schema is similar to the Network Overlap, except that we give a bonus to
-pairs that are themselves friends,
+This schema is the sum of the network overlap and direct friends schemata:
+
+$$ w_{ij} =
+  \begin{cases}
+    \frac{|\phi_i \cup \phi_j|} {|\phi_i \cap \phi_j|} + 1 & 
+      \text{if } i \in \phi_j \\
+    \frac{|\phi_i \cup \phi_j|} {|\phi_i \cap \phi_j|}      & 
+      \text{if } i \notin \phi_j
+  \end{cases} $$
+
 ```{r direct_network_overlap}

### 2015-03-27 article.Rmd
Basic document cleanup.

@@
 -----------------------------------------
-The social network data we retrieved from Facebook allow us to construct four social weights matrices, which we describe in this section.
-To aid this presentation, we must first establish some basic notation following elementary set theory.
-The set of friends of an individual $i$ is defined as $\phi_i$.
-The cardinality of this set is the number of friends, so therefore the quantity $|\phi_i \cup \phi_j|$ represents the number of friends that individuals $i$ and $j$ have in common.
+The social network data we retrieved from Facebook allow us to construct at least
+four social weights matrices, which we describe in this section. To aid this 
+presentation, we must first establish some basic notation following elementary 
+set theory. The set of friends of an individual $i$ is defined as $\phi_i$. 
+The cardinality of this set is the number of friends, so therefore the quantity 
+$|\phi_i \cup \phi_j|$ represents the number of friends that individuals $i$ 
+and $j$ have in common.
 
@@
 relationships through the API; indeed, this affected approximately 10% of cases.
-We therefore coerce this and other schema to be symmetric as appropriate.
+We therefore coerced all schemata to be symmetric as appropriate.
 
-As a final note on the schema development, behavioral inference on and
+```{r force_symmetric}
+# still need to implement
+```
+
+As a final note on the schemata development, behavioral inference on and
 estimation of the models is simplified by using a row-standardized matrix, where
-the rows sum to unity(LeSage & Pace 2009, p.85). We follow this convention
+the rows sum to unity(LeSage & Pace 2009, p.85). We follow this convention
 in all cases.
+
+```{r save_as_matrices}
+
+```
 

### 2015-04-10 article/article.Rmd
repository cleanup, added old word docs.

--- 2015-03-27 article.Rmd
@@
 a special case of social interaction, or as a proxy that may be more easily
-observed in data.
+observed in data. In other words, the spatial relationships that many authors
+have observed are often substitutes for implied social relationships.
 
@@
 directly from Facebook (i.e. not subject to a respondent recollection ability) 
-  2. Whole Networks: Facebook.com contains complete network data of a
+  2. Whole Networks: Facebook contains complete network data of a
 sociocentric nature, as opposed to an egocentric nature, which reflects the
@@
 translated to geospatial
+
+Tie Strength
+---------------------------------------
 
@@
 
+```{r number_trips}
+#calculate number of user trips
+number_trips <- trips %>%
+  group_by(id) %>%
+  summarise(number_trips = n())
+
+user <- user %>%
+  left_join(number_trips)
+```
+
 Because the survey program collects Facebook profile and friends list data from
@@
 
-Facebook locational data (i.e., hometown and current city location) is not
-geocoded, and therefore required some post-processing prior to any analyses.
-Facebook location data were post-processed using R statistical programming
-language and ArcGIS 10.  The study analyzed respondents’ friends who do not
-consider Atlanta, Georgia as their hometown. Friends’ hometown cities were
-geocoded as point locations. To associate cities as within (or outside of) the
-Atlanta metropolitan region, circular buffers were created (using
-Hartsfield-Jackson Atlanta International airport as the centroid) at 25 mile
-radius increments from 25 miles to 150 miles. The 6 resulting hometown buffer
-regions were then associated with the geocoded city locations, and thus the
-number of friends with hometown city locations within each buffer region were
-determined. Only data from respondents that completed the entire survey (i.e.
-the air travel diary, socio-demographic section, and the Facebook data
-collection) were considered for the analysis.
+***discussion of why we care about how far friends are from Atlanta***
+
+```{r distance_from_atlanta}
+cities <- read.csv("data/PRIVATEDATA/geocode/geocoded_datatables/allcities.csv")%>%
+  tbl_df()
+
+```
 

### 2015-04-10 rscripts/datacleaner.R
cleaned trip and friend file with user file.

@@ -1,17 +1,18 @@
 # Load and do basic data munging on fraternity surveys.
 library(dplyr)
+library(readr)
 library(data.table)
 library(igraph)
 
-# Load user data and trim to useable fields
-user <- read.csv("../data/PRIVATEDATA/user.csv", stringsAsFactors = FALSE) %>%
+# Clean user table
+user <- tbl_df(read_csv("../data/PRIVATEDATA/user.csv")) %>%
   select(
-    id, fb_id, zip, location_from, lives_in, age, gender, race,
-    friend_list_size, collected_friends_size
-    ) %>%
+    id, fb_id, zip, location_from, lives_in, age, gender, race, 
+    collected_friends_size, fraternity_id
+    )  %>%
   setnames(
-    c("id", "fbid", "zip", "from", "at", "age", "sex", "race", "friends", 
-      "collected")
+    c("id", "fbid", "zip", "from", "at", "age", "sex", "race", 
+      "collected", "fraternity_id")
     ) 
 
 # useable records must have friends collected
@@ -21,12 +22,37 @@ user <- user %>%
 
 # make sure numeric cols are num, character are chr
 user <- user %>%
-  mutate_each(funs(as.numeric(.)), age, friends, collected) %>%
+  mutate_each(funs(as.numeric(.)), age,  collected) %>%
   mutate(zip = sprintf("%05d", zip))
+
+# attach fraternity information
+user <- user %>%
+  left_join(read.csv("../data/fraternities.csv", colClasses = "character"))
+
 
 saveRDS(user, file = "../data/clean/user.rds")
 
-# calculate trips for each user
+
+# Clean trip data
+trips <- read_csv("../data/PRIVATEDATA/trip.csv") %>%
+  tbl_df()  %>%
+  select(
+    id = user_id, 
+    purpose, 
+    origin_airport = origin_airport_code, 
+    destination_airport = destination_airport_code) 
+
+saveRDS(trips, file = "../data/clean/trips.rds")
+
+# Clean friends data
+friends <- read_csv("../data/PRIVATEDATA/user_friends_info.csv") %>%
+  tbl_df() %>%
+  transmute(
+    fbid = fb_id,
+    id = user_id,
+    from = tolower(location_from),
+    at = tolower(lives_in)
+  )
 
 
 saveRDS(friends, file = "../data/clean/friends.rds")

### 2015-04-10 article/article.Rmd
Created graph and adjacency matrices from data.

@@
 library(dplyr, warn.conflicts = FALSE) # super data wrangling
+library(tidyr)
 library(igraph) # graph analysis
@@
 
-```{r load_user}
-trips <- readRDS("data/clean/trips.rds")
-user <- readRDS("data/clean/user.rds")
-friends <- readRDS("data/clean/friends.rds")
+```{r load_user, warning=FALSE}
+source("../rscripts/datacleaner.R")
 ```
@@
 
-```{r force_symmetric}
-# still need to implement
+```{r save_as_matrices}
+# make graph
+user_graph <- graph.data.frame(friend_connections)
+
+# export graph edges as adjacency matrix
+matrix_list <- list()
+for(i in names(friend_connections)[-(1:2)]){
+  matrix_list[[i]] <- get.adjacency(user_graph, type = "both", i)
+
+  # TODO force symmetric
+}
 ```
@@
 in all cases.
-
-```{r save_as_matrices}
-
-```
 

### 2015-04-10 article/article.Rmd
Added reference and cleaned up document.

@@
     fig_caption: yes
+    theme: journal
     number_sections: yes
@@
 
-```{r distance_from_atlanta}
+```{r distance_from_atlanta, eval=FALSE}
 cities <- read.csv("data/PRIVATEDATA/geocode/geocoded_datatables/allcities.csv")%>%
@@
 **Direct Friendship**
-This schema defines a simple binary matrix where a cell is nonzero if  and  are
-friends. Specifically,
+This schema defines a simple binary matrix where a cell is nonzero if $i$ and 
+$j$ are friends. Specifically,
 
@@
 **Mutual Friendship**
-This schema defines a matrix where the cell value is the number of friends  and
-have in common (as in Gilbert & Karahalios 2009; He et al. 2012),
+This schema defines a matrix where the cell value is the number of friends $i$
+and $j$ have in common [as in @Gilbert2009; @He2012],
 
@@
 This schema defines a matrix where the cell value is the relative overlap of
-each pair’s social network (as in Adamic & Adar 2003),
+each pair’s social network [as in @Adamic2003],
 

### 2015-04-24 article/article.Rmd
Added more math formatting.

@@
 socioeconomic variables. The Poisson count model estimates the effect of
-dependent variables  on the conditional mean number of trips, . We consider
-social interaction in this model by suggesting that  (where  is the row of  that
-contains the weights for the neighbors of ). Our goal in this analysis is to
-recover a weighting scheme  that maximizes the model likelihood, and use this
-model to examine the inferential error that could result from ignoring social
-interdependence and correlation in such model.
+dependent variables $X_i$ on the conditional mean number of 
+trips, $\lambda_i = e^{X_i\beta}$. This mean in turn determines the probability 
+of a discrete number of event occurrences,
+
+$$P(y_i = y | X) = \frac{\lambda_i e^{-\lambda_i}}{y!}$$
+
+We consider the influence of social interaction in this model by suggesting
+
+$$\lambda_i = e^{\rho W_i y + X_i\beta + W_iX_i \gamma}$$.
+
+If the $\rho$ and $\gamma$ parameters are statistically significant, this 
+indicates air travel is *socially dependent*, or that it is influenced by factors
+that are socially dependent. Further, if the parameters on this model are 
+significantly different from the non-social model, it implies that a non-social
+model may return biased parameters and/or inaccurate inferences.
+
+### Software
 

### 2015-04-24 article/article.Rmd
Added descriptive table to document.

@@
 
-```{r load_user, warning=FALSE}
+```{r load_user, warning=FALSE, cache = FALSE}
 source("../rscripts/datacleaner.R")
@@
 user <- user %>%
-  left_join(number_trips)
+  filter(!is.na(collected)) %>%
+  left_join(number_trips) %>%
+  mutate(number_trips = ifelse(is.na(number_trips), 0, number_trips))
 ```
@@
   tbl_df()
+
+```
+
+```{r survey_summary, echo = FALSE}
+us <- user %>%
+  summarise(n = n(),
+            min_age = min(age, na.rm = TRUE),
+            max_age = max(age, na.rm = TRUE),
+            med_age = median(age, na.rm = TRUE),
+            n_other = sum(race == "Other"),
+            n_zero = sum(number_trips == 0),
+            n_one = sum(number_trips == 1),
+            n_two = sum(number_trips == 2),
+            n_three = sum(number_trips == 3),
+            n_fourplus = sum(number_trips >= 4)
+            )
+```
+
+```{r summary_table, echo = FALSE}
+t <- data_frame(
+  "Variable" = c("Users:", "Responses", "Used records", 
+                 "Minimum respondent age", "Median age", "Minorities",
+                 "Trips:", "Zero trips", "One trip", "Two trips", 
+                 "Three trips", "Four or more trips",
+                 "Friends:",
+                 "Total friends collected", "Unique friends",
+                 "Unique hometowns", "Unique current locations"),
+  " " = c("", n_collected, us$n, 
+          us$min_age, us$med_age,  us$n_other, 
+          "", us$n_zero, us$n_one, us$n_two, us$n_three, us$n_fourplus,
+          "",
+          nrow(friends), length(unique(friends$fbid)),
+          length(unique(friends$from)), length(unique(friends$at)))
+)
+kable(t)
 

### 2015-04-24 rscripts/datacleaner.R

@@ -15,10 +15,15 @@ user <- tbl_df(read_csv("../data/PRIVATEDATA/user.csv")) %>%
       "collected", "fraternity_id")
     ) 
 
+
 # useable records must have friends collected
+n_collected <- nrow(user)
 user <- user %>%
   mutate_each(funs(ifelse( . == "NULL", NA, .))) %>%
-  filter(collected > 0) 
+  filter(id > 147) %>%
+  mutate(
+    race = ifelse(race != "White" | is.na(race), "Other", "White")
+  )
 
 # make sure numeric cols are num, character are chr
 user <- user %>%

### 2015-04-24 article/article.Rmd
Forced adjacency matrices to be symmetric.

@@
 for(i in names(friend_connections)[-(1:2)]){
-  matrix_list[[i]] <- get.adjacency(user_graph, type = "both", i)
-
-  # TODO force symmetric
+  mat <- as.matrix(get.adjacency(user_graph, type = "both", i))
+  # force matrix to be symmetric
+  matrix_list[[i]] <- pmax(mat, t(mat))
 }

### 2015-05-23 article/article.Rmd
Added code to calculate friends within an arbitrary radius.

@@
 ```{r distance_from_atlanta, eval=FALSE}
-cities <- read.csv("data/PRIVATEDATA/geocode/geocoded_datatables/allcities.csv")%>%
-  tbl_df()
+# how many friends are within different distances?
+friends_distance <- friends %>%
+  group_by(id) %>%
+  summarise(
+    friends_win_10 = sum(miles_from_atl < 10, na.rm = TRUE),
+    friends_win_25 = sum(miles_from_atl < 25, na.rm = TRUE),
+    friends_win_50 = sum(miles_from_atl < 50, na.rm = TRUE),
+    friends_win_100 = sum(miles_from_atl < 100, na.rm = TRUE)
+    )
 
+user <- user %>%
+  left_join(friends_distance)
 ```
+
+A descriptive summary of the data is given in Table .
 

### 2015-05-23 rscripts/datacleaner.R

@@ -4,7 +4,7 @@ library(readr)
 library(data.table)
 library(igraph)
 
-# Clean user table
+# Clean user table -----
 user <- tbl_df(read_csv("../data/PRIVATEDATA/user.csv")) %>%
   select(
     id, fb_id, zip, location_from, lives_in, age, gender, race, 
@@ -13,32 +13,31 @@ user <- tbl_df(read_csv("../data/PRIVATEDATA/user.csv")) %>%
   setnames(
     c("id", "fbid", "zip", "from", "at", "age", "sex", "race", 
       "collected", "fraternity_id")
-    ) 
+    ) %>%
 
-
-# useable records must have friends collected
-n_collected <- nrow(user)
-user <- user %>%
+  # useable records must have friends collected
   mutate_each(funs(ifelse( . == "NULL", NA, .))) %>%
   filter(id > 147) %>%
+
+  # variable cleanup
   mutate(
-    race = ifelse(race != "White" | is.na(race), "Other", "White")
-  )
-
-# make sure numeric cols are num, character are chr
-user <- user %>%
+    race = ifelse(race != "White" | is.na(race), "Other", "White"),
+    from = tolower(from),
+    at = tolower(at),
+    zip = sprintf("%05d", zip) 
+  ) %>%
   mutate_each(funs(as.numeric(.)), age,  collected) %>%
-  mutate(zip = sprintf("%05d", zip))
-
-# attach fraternity information
-user <- user %>%
+  
+  # attach fraternity information
   left_join(read.csv("../data/fraternities.csv", colClasses = "character"))
+  
 
 
-saveRDS(user, file = "../data/clean/user.rds")
+
+#saveRDS(user, file = "../data/clean/user.rds")
 
 
-# Clean trip data
+# Clean trip data ------
 trips <- read_csv("../data/PRIVATEDATA/trip.csv") %>%
   tbl_df()  %>%
   select(
@@ -47,9 +46,30 @@ trips <- read_csv("../data/PRIVATEDATA/trip.csv") %>%
     origin_airport = origin_airport_code, 
     destination_airport = destination_airport_code) 
 
-saveRDS(trips, file = "../data/clean/trips.rds")
+#saveRDS(trips, file = "../data/clean/trips.rds")
 
-# Clean friends data
+# calculate how close friend's cities are to Atlanta -----
+# distance from Atlanta 
[…]

### 2017-07-21 article/article.Rmd
Update packages and modernize to restart

@@
 library(knitr) # nice output tables
+library(here)
 
@@
 opts_chunk$set(
-  cache = TRUE, cache.path = "./.cache/",
+  cache = TRUE, cache.path = ".cache/",
   echo = TRUE, 
-  fig.keep = TRUE, fig.path = "./.figs/"
+  fig.keep = TRUE, fig.path = ".figs/"
 )
 
-library(dplyr, warn.conflicts = FALSE) # super data wrangling
-library(tidyr)
-library(igraph) # graph analysis
-library(networkD3) # interactive graph visualizer
 ```
@@
 ```{r load_user, warning=FALSE, cache = FALSE}
-source("../rscripts/datacleaner.R")
+source(here("rscripts/datacleaner.R"))
 ```
@@
 # make graph
-user_graph <- graph.data.frame(friend_connections)
+user_graph <- igraph::graph.data.frame(friend_connections)
 
@@
 for(i in names(friend_connections)[-(1:2)]){
-  mat <- as.matrix(get.adjacency(user_graph, type = "both", i))
+  mat <- as.matrix(igraph::get.adjacency(user_graph, type = "both", i))
   # force matrix to be symmetric

### 2017-07-21 rscripts/datacleaner.R

@@ -1,22 +1,17 @@
 # Load and do basic data munging on fraternity surveys.
-library(dplyr)
-library(readr)
-library(data.table)
-library(igraph)
+library(tidyverse)
+library(here)
+
 
 # Clean user table -----
-user <- tbl_df(read_csv("../data/PRIVATEDATA/user.csv")) %>%
+user <- read_csv(here("data/PRIVATEDATA/user.csv")) %>%
   select(
-    id, fb_id, zip, location_from, lives_in, age, gender, race, 
-    collected_friends_size, fraternity_id
-    )  %>%
-  setnames(
-    c("id", "fbid", "zip", "from", "at", "age", "sex", "race", 
-      "collected", "fraternity_id")
-    ) %>%
+    id, fbid = fb_id, zip, from = location_from, at = lives_in, age, gender, race, 
+    collected = collected_friends_size, fraternity_id
+  ) %>%
 
   # useable records must have friends collected
-  mutate_each(funs(ifelse( . == "NULL", NA, .))) %>%
+  mutate_all(funs(ifelse( . == "NULL", NA, .))) %>%
   filter(id > 147) %>%
 
   # variable cleanup
@@ -26,19 +21,17 @@ user <- tbl_df(read_csv("../data/PRIVATEDATA/user.csv")) %>%
     at = tolower(at),
     zip = sprintf("%05d", zip) 
   ) %>%
-  mutate_each(funs(as.numeric(.)), age,  collected) %>%
+  mutate_at(vars(age, collected), funs(as.numeric(.))) %>%
   
   # attach fraternity information
-  left_join(read.csv("../data/fraternities.csv", colClasses = "character"))
+  left_join(read_csv(here("data/fraternities.csv"), col_types = "ccc"))
+
+n_collected <- nrow(user)
   
-
-
-
-#saveRDS(user, file = "../data/clean/user.rds")
 
 
 # Clean trip data ------
-trips <- read_csv("../data/PRIVATEDATA/trip.csv") %>%
+trips <- read_csv(here("data/PRIVATEDATA/trip.csv")) %>%
   tbl_df()  %>%
   select(
     id = user_id, 
@@ -58,7 +51,7 @@ get_miles_from_atl <- function (long2, lat2) {
 }
 
 # coordinates for the friends' cities
-cities <- read_csv("../data/PRIVATEDATA/geocode/geocoded_datatables/allcities.csv") %>%
[…]
